<<<<<<< HEAD
---
title: "Practical Machine Learning Project"
author: "Romaine Neungang"
date: "5 June 2017"
output:
  html_document: default
---

## Course Project Instruction  
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).  

## Data
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.  

## What you should submit
The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.   

## Reproducibility  
Libraries Needed
```{r library, message=FALSE, warning=FALSE, echo=TRUE}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(arm)
library(kernlab)
```

## Getting Data and Partioning the training set into two
```{r data, echo=TRUE}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testing <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))
all.equal(colnames(testing)[1:length(colnames(testing))-1], colnames(training)[1:length(colnames(training))-1])
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
myTraining <- training[inTrain, ]  
myTesting <- training[-inTrain, ]
dim(myTraining); dim(myTesting)
```
## cleaning and pre-processing
```{r cleaning, echo=TRUE}
myDataNZV <- nearZeroVar(myTraining, saveMetrics=TRUE)
myTraining <- myTraining[, !myDataNZV$nzv]
# Variables with more than 60% missing values are removed
toberem <- sapply(colnames(myTraining), function(x) if(sum(is.na(myTraining[, x])) > 0.60*nrow(myTraining))    {return(TRUE)
}else{
return(FALSE)
}
)
myTraining <- myTraining[, !toberem]
```
Variables related with data acquisition (like: id, timestamps, individuals' names, etc.) are not suitable to be used in prediction and are removed
```{r, echo=TRUE}
myTraining <- myTraining[, -(1:6)]
```
Correlation analysis:
```{r, echo=TRUE}
Hcorr <- caret::findCorrelation(cor(myTraining[, -53]), cutoff=0.8)
names(myTraining)[Hcorr]
```
Many variables are highly correlated. PCA will be used in the pre-processing. After the data cleaning the variables selected to specify the model are:
```{r, echo=TRUE}
names(myTraining)
```
## Model Specification and Cross Validation
In order to avoid overfitting and to reduce out of sample errors, TrainControl is used to perform 7-fold cross validation.
Six models are estimated: Random forest, Support Vector Machine (both radial and linear), a Neural net, a Bayes Generalized linear model and a Logit Boosted model.
```{r , warning=FALSE, message=FALSE, results="hide"}
tc <- trainControl(method = "cv", number = 7, verboseIter=FALSE , preProcOptions="pca", allowParallel=TRUE)
rf <- train(classe ~ ., data = myTraining, method = "rf", trControl= tc)
svmr <- train(classe ~ ., data = myTraining, method = "svmRadial", trControl= tc)
NN <- train(classe ~ ., data = myTraining, method = "nnet", trControl= tc, verbose=FALSE)
svml <- train(classe ~ ., data = myTraining, method = "svmLinear", trControl= tc)
bayesglm <- train(classe ~ ., data = myTraining, method = "bayesglm", trControl= tc)
logitboost <- train(classe ~ ., data = myTraining, method = "LogitBoost", trControl= tc)
```
Accuracy comparison
```{r, echo=TRUE}
model <- c("Random Forest", "SVM (radial)","LogitBoost","SVM (linear)","Neural Net", "Bayes GLM")
Accuracy <- c(max(rf$results$Accuracy),
        max(svmr$results$Accuracy),
        max(logitboost$results$Accuracy),
        max(svml$results$Accuracy),
        max(NN$results$Accuracy),
        max(bayesglm$results$Accuracy))
        
Kappa <- c(max(rf$results$Kappa),
        max(svmr$results$Kappa),
        max(logitboost$results$Kappa),
        max(svml$results$Kappa),
        max(NN$results$Kappa),
        max(bayesglm$results$Kappa))  
performance <- cbind(model,Accuracy,Kappa)
knitr::kable(performance)
```
Random forest and SVM(radial) provide the best results and will provide the predictions for the submission. Even if the Out of sample error cannot be estimated exactly, the in-sample error obtained through cross-validation is calculated over different test sets and should provide a better estimate of out-of sample error with respect to the case of no cross-validation.

Prediction of "classe" variable for the testing set
```{r, echo=TRUE}
rfPred <- predict(rf, testing)
svmrPred <- predict(svmr, testing)
```
Checking if the models give same predictions
```{r, echo=TRUE}
prediction <- data.frame(cbind(rfPred, svmrPred))
prediction$same <- with(prediction, rfPred == svmrPred)
colnames(prediction) <- c("Random Forest", "SVM (radial)", "Same Prediction")

knitr::kable(prediction)
```
## Submission to Coursera
Generation of the files to be submitted is made through the provided function
```{r, echo=TRUE}
pml_write_files = function(x){
        n = length(x)
        for(i in 1:n){
                filename = paste0("problem_id_",i,".txt")
                write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
        }
}

pml_write_files(rfPred)

```

## Conclusions

The random forest model provides an outstanding accuracy and, accordingly, the predictions for the test set were correct in 100% of the cases.


=======
---
title: "Practical Machine Learning Project"
author: "Romaine Neungang"
date: "5 June 2017"
output:
  html_document: default
---

## Course Project Instruction  
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).  

## Data
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.  

## What you should submit
The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.   

## Reproducibility  
Libraries Needed
```{r library, message=FALSE, warning=FALSE, echo=TRUE}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(arm)
library(kernlab)
```

## Getting Data and Partioning the training set into two
```{r data, echo=TRUE}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testing <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))
all.equal(colnames(testing)[1:length(colnames(testing))-1], colnames(training)[1:length(colnames(training))-1])
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
myTraining <- training[inTrain, ]  
myTesting <- training[-inTrain, ]
dim(myTraining); dim(myTesting)
```
## cleaning and pre-processing
```{r cleaning, echo=TRUE}
myDataNZV <- nearZeroVar(myTraining, saveMetrics=TRUE)
myTraining <- myTraining[, !myDataNZV$nzv]
# Variables with more than 60% missing values are removed
toberem <- sapply(colnames(myTraining), function(x) if(sum(is.na(myTraining[, x])) > 0.60*nrow(myTraining))    {return(TRUE)
}else{
return(FALSE)
}
)
myTraining <- myTraining[, !toberem]
```
Variables related with data acquisition (like: id, timestamps, individuals' names, etc.) are not suitable to be used in prediction and are removed
```{r, echo=TRUE}
myTraining <- myTraining[, -(1:6)]
```
Correlation analysis:
```{r, echo=TRUE}
Hcorr <- caret::findCorrelation(cor(myTraining[, -53]), cutoff=0.8)
names(myTraining)[Hcorr]
```
Many variables are highly correlated. PCA will be used in the pre-processing. After the data cleaning the variables selected to specify the model are:
```{r, echo=TRUE}
names(myTraining)
```
## Model Specification and Cross Validation
In order to avoid overfitting and to reduce out of sample errors, TrainControl is used to perform 7-fold cross validation.
Six models are estimated: Random forest, Support Vector Machine (both radial and linear), a Neural net, a Bayes Generalized linear model and a Logit Boosted model.
```{r , warning=FALSE, message=FALSE, results="hide"}
tc <- trainControl(method = "cv", number = 7, verboseIter=FALSE , preProcOptions="pca", allowParallel=TRUE)
rf <- train(classe ~ ., data = myTraining, method = "rf", trControl= tc)
svmr <- train(classe ~ ., data = myTraining, method = "svmRadial", trControl= tc)
NN <- train(classe ~ ., data = myTraining, method = "nnet", trControl= tc, verbose=FALSE)
svml <- train(classe ~ ., data = myTraining, method = "svmLinear", trControl= tc)
bayesglm <- train(classe ~ ., data = myTraining, method = "bayesglm", trControl= tc)
logitboost <- train(classe ~ ., data = myTraining, method = "LogitBoost", trControl= tc)
```
Accuracy comparision
```{r, echo=TRUE}
model <- c("Random Forest", "SVM (radial)","LogitBoost","SVM (linear)","Neural Net", "Bayes GLM")
Accuracy <- c(max(rf$results$Accuracy),
        max(svmr$results$Accuracy),
        max(logitboost$results$Accuracy),
        max(svml$results$Accuracy),
        max(NN$results$Accuracy),
        max(bayesglm$results$Accuracy))
        
Kappa <- c(max(rf$results$Kappa),
        max(svmr$results$Kappa),
        max(logitboost$results$Kappa),
        max(svml$results$Kappa),
        max(NN$results$Kappa),
        max(bayesglm$results$Kappa))  
performance <- cbind(model,Accuracy,Kappa)
knitr::kable(performance)
```
Random forest and SVM(radial) provide the best results and will provide the predictions for the submission. Even if the Out of sample error cannot be estimated exactly, the in-sample error obtained through cross-validation is calculated over different test sets and should provide a better estimate of out-of sample error with respect to the case of no cross-validation.

Prediction of "classe" variable for the testing set
```{r, echo=TRUE}
rfPred <- predict(rf, testing)
svmrPred <- predict(svmr, testing)
```
Checking if the models give same predictions
```{r, echo=TRUE}
prediction <- data.frame(cbind(rfPred, svmrPred))
prediction$same <- with(prediction, rfPred == svmrPred)
colnames(prediction) <- c("Random Forest", "SVM (radial)", "Same Prediction")

knitr::kable(prediction)
```
Generation of the files to be submitted is made through the provided function
```{r, echo=TRUE}
pml_write_files = function(x){
        n = length(x)
        for(i in 1:n){
                filename = paste0("problem_id_",i,".txt")
                write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
        }
}

pml_write_files(rfPred)

```

## Conclusions

The random forest model provides an outstanding accuracy and, accordingly, the predictions for the test set were correct in 90% of the cases.


>>>>>>> 884a10356254095efd247e041da5d06a2e4109ea
